Parameters estimation via HOSVD

### 2024-10-22 Code/R/Research/HOSVD parameters estimate.Rmd
Different rate cases and singular values plots

---
title: "Parameters estimation via HOSVD"
output: html_document
---
```{r setup, include=FALSE}
library(Rssa)
library(rTensor)
library(ggplot2) 
library(dplyr)
library(tidyr)
library(waldo)
```

@@

```{r hosvd-mod}
fnorm_complex <- function(tnsr) {
  sqrt(sum(abs(tnsr@data) ^ 2))
}

@@

@@

```{r 1d-comparison-for-sds, results='hide'}
len.out <- 16
sds <- seq(from = 0.01, to = 0.06, length.out = len.out)
set.seed(10)

n <- 25
L <- 15
I <- 14; L.tens <- 8
rates.true <- c(-0.01, -0.02)
freqs.true <- c(0.2, 0.22)
signal <- exp((rates.true[1] + 2i * pi * freqs.true[1]) * 0:(n - 1)) +
  exp((rates.true[2] + 2i * pi * freqs.true[2]) * 0:(n - 1))
R <- 1000

@@

    par.ssa <- parestimate(x.ssa, groups = list(1:2))
    pars.ssa$rate1[i] <- max(par.ssa$rates[1], par.ssa$rates[2])
    pars.ssa$rate2[i] <- min(par.ssa$rates[1], par.ssa$rates[2])
    pars.ssa$freq1[i] <- min(par.ssa$frequencies[1], par.ssa$frequencies[2])
    pars.ssa$freq2[i] <- max(par.ssa$frequencies[1], par.ssa$frequencies[2])

    par.tssa <- tens_esprit(x, I, L.tens, groups = list(1:2), status = FALSE)
    pars.tssa$rate1[i] <- max(par.tssa[[1]]$rates[1], par.tssa[[1]]$rates[2])
    pars.tssa$rate2[i] <- min(par.tssa[[1]]$rates[1], par.tssa[[1]]$rates[2])
    pars.tssa$freq1[i] <- min(par.tssa[[1]]$frequencies[1], par.tssa[[1]]$frequencies[2])
    pars.tssa$freq2[i] <- max(par.tssa[[1]]$frequencies[1], par.tssa[[1]]$frequencies[2])

@@

```{r plot-and-save}
plot.save <- function(plt, name, format, ...) {
  print(plt)
  format(name, ...)
  print(plt)
  dev.off()
}
```

@@

plot.save(rate1.plot,
          "./img/rate1_mini.pdf",
          pdf,
          width = 9,
          height = 6)
```

@@

plot.save(rate2.plot,
          "./img/rate2_mini.pdf",
          pdf,
          width = 9,
          height = 6)
```

@@

plot.save(freq1.plot,
          "./img/freq1_mini.pdf",
          pdf,
          width = 9,
          height = 6)
```
```{r}
freq2.df <- freq2.rrmse |> as.data.frame() |> mutate(sd = sds) |>
  pivot_longer(!sd, names_to = "type", values_to = "RRMSE")
freq2.plot <- freq2.df |> ggplot() + geom_line(aes(
  x = sd,
  y = RRMSE,
  colour = type,
  linetype = type
), lwd = 1.2)

plot.save(freq2.plot,
          "./img/freq2_mini.pdf",
          pdf,
          width = 9,
          height = 6)
```

@@

rates.true <- c(-0.01, -0.02)
freqs.true <- c(0.2, 0.22)
coef <- matrix(CCSWGN(2 * Q), ncol = 2)

signal <- exp((rates.true[1] + 2i * pi * freqs.true[1]) * 0:(n - 1)) %o% coef[, 1] + 
  exp((rates.true[2] + 2i * pi * freqs.true[2]) * 0:(n - 1)) %o% coef[, 2]
R <- 1000

@@

@@

    par.ssa <- cmesprit(x, L = L, groups = list(1:2))
    pars.ssa$rate1[i] <- max(par.ssa[[1]]$rates[1], par.ssa[[1]]$rates[2])
    pars.ssa$rate2[i] <- min(par.ssa[[1]]$rates[1], par.ssa[[1]]$rates[2])
    pars.ssa$freq1[i] <- min(par.ssa[[1]]$frequencies[1], par.ssa[[1]]$frequencies[2])
    pars.ssa$freq2[i] <- max(par.ssa[[1]]$frequencies[1], par.ssa[[1]]$frequencies[2])

    par.tssa <- tens_esprit(x, L.tens, kind = "HO-MSSA", est_dim = 1, groups = list(1:2), status = FALSE)
    pars.tssa$rate1[i] <- max(par.tssa[[1]]$rates[1], par.tssa[[1]]$rates[2])
    pars.tssa$rate2[i] <- min(par.tssa[[1]]$rates[1], par.tssa[[1]]$rates[2])
    pars.tssa$freq1[i] <- min(par.tssa[[1]]$frequencies[1], par.tssa[[1]]$frequencies[2])
    pars.tssa$freq2[i] <- max(par.tssa[[1]]$frequencies[1], par.tssa[[1]]$frequencies[2])

@@

plot.save(rate1.plot,
          "./img/rate1_nd.pdf",
          pdf,
          width = 9,
          height = 6)
```

@@

plot.save(rate2.plot,
          "./img/rate2_nd.pdf",
          pdf,
          width = 9,
          height = 6)
```

@@

plot.save(freq1.plot,
          "./img/freq1_nd.pdf",
          pdf,
          width = 9,
          height = 6)
```
```{r}
freq2.df <- freq2.rrmse |> as.data.frame() |> mutate(sd = sds) |>
  pivot_longer(!sd, names_to = "type", values_to = "RRMSE")
freq2.plot <- freq2.df |> ggplot() + geom_line(aes(
  x = sd,
  y = RRMSE,
  colour = type,
  linetype = type
), lwd = 1.2)

plot.save(freq2.plot,
          "./img/freq2_nd.pdf",
          pdf,
          width = 9,
          height = 6)
```

>>>>

---
title: "Parameters estimation via HOSVD"
output: html_document
---
```{r setup, include=FALSE}
library(Rssa)
library(rTensor)
library(ggplot2)
library(gridExtra)
library(factoextra)
library(dplyr)
library(tidyr)
library(waldo)
```

@@

```{r hosvd-mod}
fnorm_complex <- function(x) {
  sqrt(sum(abs(x) ^ 2))
}

setMethod("fnorm_complex", "Tensor", function(tnsr) {
  sqrt(sum(abs(tnsr@data) ^ 2))
})

@@

```{r plot-and-save}
plot.save <- function(plt, name, format, ..., save = TRUE) {
  print(plt)
  if (save){
    format(name, ...)
    print(plt)
    dev.off()
  }
}
save = FALSE
```

@@

```{r}
len.out <- 16
sds <- seq(from = 0, to = 0.2, length.out = len.out)

set.seed(10)

n <- 25
L <- 15
I <- 14; L.tens <- 8

# no rates
rates.true <- c(0, 0)
# both rates -0.01
# rates.true <- c(-0.01, -0.01)
# both rates -0.02
# rates.true <- c(-0.02, -0.02)
# rates from paper
# rates.true <- c(-0.01, -0.02)

freqs.true <- c(0.2, 0.22)
signal <- exp((rates.true[1] + 2i * pi * freqs.true[1]) * 0:(n - 1)) +
  exp((rates.true[2] + 2i * pi * freqs.true[2]) * 0:(n - 1))

R <- 100
l.range <- 2:4

lambda.mean <- list(ssa = matrix(numeric(length(l.range) * len.out), nrow = length(l.range)), 
                    tssa = matrix(numeric(length(l.range) * len.out), nrow = length(l.range)))
colnames(lambda.mean$ssa) <- round(sds, 2)
colnames(lambda.mean$tssa) <- round(sds, 2)
rownames(lambda.mean$ssa) <- l.range
rownames(lambda.mean$tssa) <- l.range

for (j in seq_along(sds)) {
  sd <- sds[j]

  lambda.ssa <- matrix(numeric(length(l.range) * R), nrow = length(l.range))
  lambda.tssa <- matrix(numeric(length(l.range) * R), nrow = length(l.range))

    lambda.ssa[,i] <- svd(Rssa::hankel(x, L))$d[l.range]

    Z <- hosvd_mod(tens3(x, I, L.tens), status = FALSE)$Z[,, l.range, drop = FALSE]@data
    lambda.tssa[,i] <- apply(Z, 3, fnorm_complex)

  lambda.mean$ssa[,j] <- rowMeans(lambda.ssa)
  lambda.mean$tssa[,j] <- rowMeans(lambda.tssa)
}
close(pb)
```

```{r}
scree.plots.ssa <- list()
scree.plots.tssa <- list()

for (j in 1:len.out) {
  df <- data.frame(l_ssa = lambda.mean$ssa[, j], l_tssa = lambda.mean$tssa[, j])

  scree.plots.ssa[[j]] <- df |> ggplot(aes(x = l.range, y = l_ssa)) +
    geom_point(size = 4) + geom_line() +
    labs(title = paste0("SSA, sd = ", round(sds[j], 2)),
         x = "Index", y = "Lambda") +
    ylim(c(min(lambda.mean$ssa), max(lambda.mean$ssa)))

  scree.plots.tssa[[j]] <- df |> ggplot(aes(x = l.range, y = l_tssa)) +
    geom_point(size = 4) + geom_line() +
    labs(title = paste0("HO-SSA, sd = ", round(sds[j], 2)),
         x = "Index", y = "Lambda") +
    ylim(c(min(lambda.mean$tssa), max(lambda.mean$tssa)))
}

ggsave("./img/Scree_plots_ssa.pdf",
       marrangeGrob(grobs = scree.plots.ssa, ncol = 2, nrow = 3,
                    layout_matrix = matrix(seq_len(2 * 3), nrow = 3, byrow = TRUE)),
       width = 7, height = 7)

ggsave("./img/Scree_plots_tssa.pdf",
       marrangeGrob(grobs = scree.plots.tssa, ncol = 2, nrow = 3,
                    layout_matrix = matrix(seq_len(2 * 3), nrow = 3, byrow = TRUE)),
       width = 7, height = 7)
```

```{r 1d-comparison-for-sds, results='hide'}
len.out <- 16
sds <- seq(from = 0.01, to = 0.2, length.out = len.out)
# sds <- seq(from = 0.01, to = 0.06, length.out = len.out)
set.seed(10)

n <- 25
L <- 15
I <- 14; L.tens <- 8
# no rates
# rates.true <- c(0, 0)
# both rates -0.01
# rates.true <- c(-0.01, -0.01)
# both rates -0.02
# rates.true <- c(-0.02, -0.02)
# rates from paper
rates.true <- c(-0.01, -0.02)

freqs.true <- c(0.2, 0.22)
signal <- exp((rates.true[1] + 2i * pi * freqs.true[1]) * 0:(n - 1)) +
  exp((rates.true[2] + 2i * pi * freqs.true[2]) * 0:(n - 1))

groups <- list(1:2)

R <- 500

@@

    par.ssa <- parestimate(x.ssa, groups = groups)
    pars.ssa$rate1[i] <- max(par.ssa$rates[1], par.ssa$rates[2])
    pars.ssa$rate2[i] <- min(par.ssa$rates[1], par.ssa$rates[2])
    pars.ssa$freq1[i] <- min(par.ssa$frequencies[1], par.ssa$frequencies[2])
    pars.ssa$freq2[i] <- max(par.ssa$frequencies[1], par.ssa$frequencies[2])

    par.tssa <- tens_esprit(x, I, L.tens, groups = groups, status = FALSE)
    pars.tssa$rate1[i] <- max(par.tssa[[1]]$rates[1], par.tssa[[1]]$rates[2])
    pars.tssa$rate2[i] <- min(par.tssa[[1]]$rates[1], par.tssa[[1]]$rates[2])
    pars.tssa$freq1[i] <- min(par.tssa[[1]]$frequencies[1], par.tssa[[1]]$frequencies[2])
    pars.tssa$freq2[i] <- max(par.tssa[[1]]$frequencies[1], par.tssa[[1]]$frequencies[2])

@@

```{r}
signal.plot.df <- data.frame(
  re_1 = Re(exp((rates.true[1] + 2i * pi * freqs.true[1]) * 0:(n - 1))), 
  im_1 = Im(exp((rates.true[1] + 2i * pi * freqs.true[1]) * 0:(n - 1))),
  re_2 = Re(exp((rates.true[2] + 2i * pi * freqs.true[2]) * 0:(n - 1))),
  im_2 = Im(exp((rates.true[2] + 2i * pi * freqs.true[2]) * 0:(n - 1))),
  index = 1:n) |> mutate(re_sum = re_1 + re_2, im_sum = im_1 + im_2)

signal.plot.df |> pivot_longer(c(re_1, re_2, re_sum), names_to = "Re_num", 
                               names_prefix = "re_", values_to = "Re") |>
  ggplot(aes(x = index)) + geom_line(aes(y = Re, col = Re_num, linetype = Re_num), lwd = 1.2)

signal.plot.df |> pivot_longer(c(im_1, im_2, im_sum), names_to = "Im_num", 
                               names_prefix = "im_", values_to = "Im") |>
  ggplot(aes(x = index)) + geom_line(aes(y = Im, col = Im_num, linetype = Im_num), lwd = 1.2)
```

@@

plot.save(rate1.plot,
          "./img/rate1.pdf",
          pdf,
          width = 9,
          height = 6, save = save)
```

@@

plot.save(rate2.plot,
          "./img/rate2.pdf",
          pdf,
          width = 9,
          height = 6, save = save)
```

@@

plot.save(freq1.plot,
          "./img/freq1.pdf",
          pdf,
          width = 9,
          height = 6, save = save)
```
```{r}
freq2.df <- freq2.rrmse |> as.data.frame() |> mutate(sd = sds) |>
  pivot_longer(!sd, names_to = "type", values_to = "RRMSE")
freq2.plot <- freq2.df |> ggplot() + geom_line(aes(
  x = sd,
  y = RRMSE,
  colour = type,
  linetype = type
), lwd = 1.2)

plot.save(freq2.plot,
          "./img/freq2.pdf",
          pdf,
          width = 9,
          height = 6, save = save)
```

@@

# no rates
rates.true <- c(0, 0)
# both rates -0.01
# rates.true <- c(-0.01, -0.01)
# both rates -0.02
# rates.true <- c(-0.02, -0.02)
# rates from paper
# rates.true <- c(-0.01, -0.02)

freqs.true <- c(0.2, 0.22)

coef <- matrix(CCSWGN(2 * Q), ncol = 2)

signal.mult <- exp((rates.true[1] + 2i * pi * freqs.true[1]) * 0:(n - 1)) %o% coef[, 1] + 
  exp((rates.true[2] + 2i * pi * freqs.true[2]) * 0:(n - 1)) %o% coef[, 2]

R <- 1000

rate1.mult.rrmse <- list(mssa = numeric(len.out), tmssa = numeric(len.out))
rate2.mult.rrmse <- list(mssa = numeric(len.out), tmssa = numeric(len.out))
freq1.mult.rrmse <- list(mssa = numeric(len.out), tmssa = numeric(len.out))
freq2.mult.rrmse <- list(mssa = numeric(len.out), tmssa = numeric(len.out))

@@

for (j in seq_along(sds)) {
  pars.mssa <- list(
    rate1 = numeric(R),
    rate2 = numeric(R),
    freq1 = numeric(R),
    freq2 = numeric(R)
  )

  pars.tmssa <- list(
    rate1 = numeric(R),
    rate2 = numeric(R),
    freq1 = numeric(R),
    freq2 = numeric(R)
  )

@@

  for (i in seq(R)) {
    noise <- CCSWGN(n, mean = 0, sd = sd)
    x <- signal.mult + noise

    par.mssa <- cmesprit(x, L = L, groups = list(1:2))
    pars.mssa$rate1[i] <- max(par.mssa[[1]]$rates[1], par.mssa[[1]]$rates[2])
    pars.mssa$rate2[i] <- min(par.mssa[[1]]$rates[1], par.mssa[[1]]$rates[2])
    pars.mssa$freq1[i] <- min(par.mssa[[1]]$frequencies[1], par.mssa[[1]]$frequencies[2])
    pars.mssa$freq2[i] <- max(par.mssa[[1]]$frequencies[1], par.mssa[[1]]$frequencies[2])

    par.tmssa <- tens_esprit(x, L.tens, kind = "HO-MSSA", est_dim = 1, groups = list(1:2), status = FALSE)
    pars.tmssa$rate1[i] <- max(par.tmssa[[1]]$rates[1], par.tmssa[[1]]$rates[2])
    pars.tmssa$rate2[i] <- min(par.tmssa[[1]]$rates[1], par.tmssa[[1]]$rates[2])
    pars.tmssa$freq1[i] <- min(par.tmssa[[1]]$frequencies[1], par.tmssa[[1]]$frequencies[2])
    pars.tmssa$freq2[i] <- max(par.tmssa[[1]]$frequencies[1], par.tmssa[[1]]$frequencies[2])

@@

  rate1.mult.rrmse$mssa[j] <- rrmse(rates.true[1], pars.mssa$rate1)
  rate1.mult.rrmse$tmssa[j] <- rrmse(rates.true[1], pars.tmssa$rate1)
  rate2.mult.rrmse$mssa[j] <- rrmse(rates.true[2], pars.mssa$rate2)
  rate2.mult.rrmse$tmssa[j] <- rrmse(rates.true[2], pars.tmssa$rate2)
  freq1.mult.rrmse$mssa[j] <- rrmse(freqs.true[1], pars.mssa$freq1)
  freq1.mult.rrmse$tmssa[j] <- rrmse(freqs.true[1], pars.tmssa$freq1)
  freq2.mult.rrmse$mssa[j] <- rrmse(freqs.true[2], pars.mssa$freq2)
  freq2.mult.rrmse$tmssa[j] <- rrmse(freqs.true[2], pars.tmssa$freq2)
}
close(pb)
```

```{r}
save = TRUE

rate1.df <- rate1.rrmse |> as.data.frame() |> mutate(sd = sds) |>
  pivot_longer(!sd, names_to = "type", values_to = "RRMSE")
rate1.plot <- rate1.df |> ggplot() +
  geom_line(aes(
    x = sd,
    y = RRMSE,
    colour = type,
    linetype = type
  ), lwd = 1.2)

plot.save(rate1.plot,
          "./img/rate1_nd.pdf",
          pdf,
          width = 9,
          height = 6, save = save)
```

@@

plot.save(rate2.plot,
          "./img/rate2_nd.pdf",
          pdf,
          width = 9,
          height = 6, save = save)
```

@@

plot.save(freq1.plot,
          "./img/freq1_nd.pdf",
          pdf,
          width = 9,
          height = 6, save = save)
```

```{r}
freq2.df <- freq2.rrmse |> as.data.frame() |> mutate(sd = sds) |>
  pivot_longer(!sd, names_to = "type", values_to = "RRMSE")
freq2.plot <- freq2.df |> ggplot() + geom_line(aes(
  x = sd,
  y = RRMSE,
  colour = type,
  linetype = type
), lwd = 1.2)

plot.save(freq2.plot,
          "./img/freq2_nd.pdf",
          pdf,
          width = 9,
          height = 6, save = save)
```
